---
title: "HW1"
output: github_document
date: "2022-09-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls()) #initialization
library(dplyr)
library(nycflights13)
library(ggplot2)
library(sjmisc)
```

### question 1

> How many flights have a missing dep_time? What other variables are missing? What might these rows represent?

```{r}

ques1 = flights %>% 
  group_by(tailnum) %>% 
  summarise(missing_dep_time = sum(is.na(dep_time)), n = n()) %>% 
  arrange(desc(missing_dep_time))

ques1_b = ques1 %>% 
  filter(missing_dep_time == 1)

nrow(ques1)
nrow(ques1_b)

other = sapply(flights, function(x) sum(is.na(x)))
la = labels(other[other > 0])
la
```
***4044 flights have missing dep_time*** \
***"dep_time"  "dep_delay" "arr_time"  "arr_delay" "tailnum"   "air_time" are missing*** \
***these rows might mean the flight have been cancelled, no dep and arr time.***

### question 2

> Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

```{r}
ques2 = flights %>% 
  mutate(dep_time = (dep_time %/% 100) * 60 + (dep_time %% 100), sched_dep_time = (sched_dep_time %/% 100) * 60 + (sched_dep_time %% 100))
tail(ques2,10)


```

### question 3

> Look at the number of canceled flights per day. Is there a pattern? Is the proportion of canceled flights related to the average delay? Use multiple dyplr operations, all on one line, concluding with ggplot(aes(x= ,y=)) + geom_point()

```{r}
#[1] "dep_time"  "dep_delay" "arr_time" 
#[4] "arr_delay" "tailnum"   "air_time" 

ques3 = flights %>% 
  mutate(cancel = (is.na(dep_time)) | (is.na(dep_delay)) | (is.na(arr_time)) | (is.na(arr_delay)) | (is.na(air_time)), delay = if_else(condition = arr_delay > 0,
                                    true = T,
                                    false = F,
                                    missing = T)) %>% 
  group_by(day) %>% 
  summarise(po = mean(cancel), sum = sum(cancel), ave_delay = mean(delay),n = n())
ques3 %>% 
  ggplot(aes(x = day)) +
    geom_point(aes(y = sum)) +
    geom_line(aes(y = po)) +
    geom_line(aes(y = ave_delay), colour = "red")

ques3 %>% 
  ggplot(aes(x = day)) +
    geom_line(aes(y = po)) +
    geom_line(aes(y = ave_delay), colour = "red")
print(ques3, n=31)
```

